---
title: "Multiscale Benchmarks"
output: html_notebook
---

```{r}
## Given an (nxd) matrix, produces a new matrix on the order of O(nd) of distinct 
## nondecreasing combinations of the original matrix
nondecreasing_matrix <- function(mat){
  n <- nrow(mat)
  d <- ncol(mat)
  idx <- rep(1, d)
  res <- list()
  while(any(idx < n)){
    ## extract current row 
    c_row <- sapply(seq(d), function(d_i){ mat[idx[d_i],d_i] })
    
    ## get index of next largest element 
    min_idx <- which.min(ifelse(idx < n, c_row, Inf))
    
    ## Increase row
    n_row <- sapply(seq(d), function(d_i){
        mat[idx[d_i], d_i]
    })
    
    ## append
    res <- append(res, list(n_row))
    
    ## increment 
    idx[min_idx] <- idx[min_idx] + 1
  }
  do.call(rbind, res)
}

## Benchmark
## Given a function that returns a Mapper object and a list of discretizations
bench_mapper <- function(mapper_f, g_card) {
  lapply(g_card, function(degree){
  
    ## Get a new mapper
    m <- mapper_f()
    
    ## Get the overlap values 
    if (is.numeric(degree)){
      ## How good of an approximation to the multiscale mapper is wanted
      g <- replicate(ncol(m$cover$filter_values), seq(0, 50, length.out = degree))
    } else if (degree == "exact"){
      filter_rng <- apply(m$cover$filter_values, 2, range)
      { filter_min <- filter_rng[1,]; filter_max <- filter_rng[2,] }
      filter_len <- diff(filter_rng)
      r_len <- filter_len/m$cover$number_intervals
      d <- ncol(m$cover$filter_values)
      g <- do.call(cbind, lapply(seq(d), function(d_i){
        as.vector(1) - (as.vector(r_len[[d_i]])/m$multiscale$filt_dist[[d_i]])
      }))
      g <- g[apply(g, 1, function(g_i) all(g_i < 0.50)),,drop = FALSE]*100
      g <- nondecreasing_matrix(g)
    }
    print("calculated g")
  
    ## Cumulative number of set / pairs need to be computed
    cum_num_updates <- apply(g, 1, function(g_i){
      print(g_i)
      
      ## Worst case -- no knowledge
      worst_sets <- prod(m$cover$number_intervals)
      worst_pairs <- choose(worst_sets, 2)
      
      ## Bounded case 
      bounded_cover <- m$cover$clone()
      bounded_cover$percent_overlap <- g_i
      bounded_pairs <- nrow(bounded_cover$level_sets_to_compare())
      
      ## Our approach
      tmp <- m$update_mapper(g_i, stats = TRUE)
      indexed_sets <- length(tmp$updated_ls)
      indexed_pairs <- nrow(tmp$updated_ls_pairs)
      
      list(
        sets = list(worst_sets, indexed_sets),
        pairs = list(worst_pairs, bounded_pairs, indexed_pairs)
      )
    })
    
    ## Accumulate the results
    set_updates <- do.call(rbind, lapply(cum_num_updates, function(lst) lst$sets))
    set_df <- data.frame(resolution=degree, apply(set_updates, 2, cumsum), percent_overlap=g)
    pair_updates <- do.call(rbind, lapply(cum_num_updates, function(lst) lst$pairs))
    pair_df <- data.frame(resolution=degree, apply(pair_updates, 2, cumsum), percent_overlap=g)
    return(list(set_res = set_df, pair_res = pair_df))
  })
}
```

```{r}
## Diabetes data from the original paper + KDE 
data("chemdiab", package = "locfit")
X <- apply(chemdiab[, 1:5], 2, function(x) scale(x, center = FALSE))
f_kde <- ks::kde(X, H = diag(5), eval.points = X, verbose = FALSE)$estimate
f_ecc <- apply(as.matrix(dist(X)), 1, max)

## Diabetes mapper
diab_mapper <- function(Z){
  return(function(){
    MapperRef$new(X = X)$
      use_cover(filter_values = matrix(Z), typename = "fixed rectangular", number_intervals = 4L, percent_overlap = 0)$
      use_distance_measure(measure = "euclidean")$
      use_clustering_algorithm(cl = "single", num_bins = 15L)$
      compute_k_skeleton(k = 1L)$
      enable_multiscale() ## enables multiscale mapper
  })
}

## Computes the mapper at the given discrete approximations. 'Exact' computes all poss
g_card <- list(10, 50, "exact")
diab_bench_kde <- do.call(rbind, bench_mapper(diab_mapper(f_kde), g_card))
diab_bench_ecc <- do.call(rbind, bench_mapper(diab_mapper(f_ecc), g_card))
saveRDS(diab_bench_kde, file = "diab_bench_kde.rds")
saveRDS(diab_bench_ecc, file = "diab_bench_ecc.rds")
```

$k$-dimensional benchmarks
```{r}
## World Values Survey data
library("Mapper")

## World Values Survey 
WV6_Data_R <- readRDS("~/WaCS/reporting/SIAGA - Mapper Parameterization/data/WV6_Data_R_v_2016_01_01.rds")
WV_US_survey <- data.table::data.table(WV6_Data_R)[V2 == 840]

## Choose some interesting dimensions to consider. Feel free to swap these in and out based on your interests! 
key_vals <- c(V4="FamilyImportance", V5="FriendImportance", V7="PoliticsImportance", V8="WorkImportance", 
              V9="ReligionImportance", V10="Happiness", V11="Health", V23="Satisfied", V24="TrustInOthers", 
              V25="ReligiousMember", V29="PoliticsMember", V30="EnvironMember", V32="CharityMember", 
              V55="ControlOverLife", V56="TakeAdvantageOf", V57="MarriedStatus", V58="NumChildren", 
              V59="FinancialSatisfaction", V60="CountryAimsFirstChoice", V61="CountryAimsSecondChoice", 
              V67="FutureWorkEmphasis", V68="FutureTechnologyEmphasis", V69="FutureRespectAuthority", 
              V84="InterestInPolitics", V95="PoliticalLeaning", V97="OwnershipOfBusiness", V98="BeliefInCompetition", 
              V101="BeliefOnWealth")

## Scale the data. This also recenters the responses to a common location. 
WV_US.scaled <- apply(WV_US_survey[,.SD, .SDcols = names(key_vals)], 2, scale)
colnames(WV_US.scaled) <- key_vals

## Extract the most statistically independent components w/ ICA, and the first 3 principle components w/ PCA.
wv_ica <- fastICA::fastICA(as.matrix(WV_US.scaled), n.comp=3, method = "C", alg.typ = "deflation", fun = "logcosh")
wv_pca <- stats::prcomp(as.matrix(WV_US.scaled), center=FALSE, rank=3) ## equiv. to wv_ica$X %*% wv_ica$K

## Use empirical distance-to-measure (DTM) to remove outliers
detect_outliers <- function(X, percentile=0.99, k=15L){
  knn <- RANN::nn2(X, k = k)
  dtm <- rowSums(knn$nn.dist)/k
  which(dtm >= quantile(dtm, c(percentile)))
}
ica_outliers <- detect_outliers(wv_ica$S, percentile = 0.95)
pca_outliers <- detect_outliers(wv_pca$x, percentile = 0.95)

## World Values Survey Mapper
wvs_mapper <- function(X, Z){
  return(function(){
    m <- Mapper:::MapperRef$new(X)$
      use_cover(filter_values=Z, type="fixed rectangular", number_intervals=5L, percent_overlap=0)$
      use_distance_measure(measure = "euclidean")
    # use_clustering_algorithm(cl = "single", num_bins = 10L)$ # For analysis 
      m$clustering_algorithm <- function(X, idx, ...){ rep(1L, length(idx)) }
      m$compute_k_skeleton(k = 1L)$
      enable_multiscale() ## enables multiscale mapper
  })
}


## Computes the mapper at the given discrete approximations. 'Exact' computes all poss
g_card <- list(10, 100, "exact")
wvs_ica_mapper_f <- wvs_mapper(as.matrix(WV_US.scaled)[-ica_outliers,], wv_ica$S[-ica_outliers,])
wvs_pca_mapper_f <- wvs_mapper(as.matrix(WV_US.scaled)[-pca_outliers,], wv_pca$x[-pca_outliers,])
wvs_bench_ica <- do.call(rbind, bench_mapper(wvs_ica_mapper_f, g_card))
wvs_bench_pca <- do.call(rbind, bench_mapper(wvs_pca_mapper_f, g_card))
saveRDS(wvs_bench_ica, file = "wvs_bench_ica.rds")
saveRDS(wvs_bench_pca, file = "wvs_bench_pca.rds")
```

Torus example data set 
```{r}
# Rejection sampling method of sampling from a torus 
# Code from "Sampling from A Manifold" by Diaconis, Holmes and Shahshahani
# http://arxiv.org/pdf/1206.6913.pdf

## Rejection-sampling of torus 
reject <- function(n=100,r=0.5,R=1) {
  xvec <- runif(n,0,2*pi)
  yvec <- runif(n,0,1/pi)
  fx <- (1+(r/R)*cos(xvec))/(2*pi)
  xvec[yvec<fx]
}

## Torus in R^3
M <- local({
  n <- 10000L*2 ## number of points to sample 
  R <- 6 ## distance from the center of interior tube to center of torus
  r <- 3 ## radius of tube
  p <- 4 ## transformed variable
  q <- reject(n, r = r, R = R) ## use runif(n, min = 0, max = 2*p) for comparison
  y <- runif(length(q), min = 0, max = 2*p)
  cbind((R + r*cos(q))*cos(y), (R + r*cos(q))*sin(y), sin(q))
})
M <- apply(M, 2, scale) # rgl::plot3d(M) to see the point cloud

## Apply a random rotation
set.seed(1234)
M.pad <- cbind(M, replicate(27L, rep(0L, nrow(M))))
QR.norm <- qr(x = replicate(30, rnorm(30)))
M_emb <- M.pad %*% QR.norm$qr
```

<!-- Loading a conda environment w/ umap and scikit installed -->
```{r, echo = FALSE, eval=TRUE}
## Load UMAP using reticulate
library("reticulate")
reticulate::use_python(normalizePath("~/miniconda3/envs/XAI/bin/python3"), required = TRUE)
reticulate::use_condaenv("XAI", required = TRUE)
# sys <- import("sys")
# cat(sys$version)
# cat("\nPath: ", head(sys$path, 1))
```

Laplacian Eigenmaps and UMAP filters
```{r}
## Build the heated graph 
M_emb.scale <- apply(M_emb, 2, scale)
torus_dist <- parallelDist::parallelDist(M_emb.scale, method = "euclidean")
sigma <- mean(apply(M_emb.scale, 2, stats::bw.SJ))
torus_kernel <- exp(-(torus_dist/(sigma*10))) ## 10 might be the smallest, 25 is too large
aff_matrix <- as.matrix(torus_kernel) ## Precompute the 'affinity' matrix
# rgl::plot3d(M_emb, col = bin_color(aff_matrix[1,]))
# rgl::texts3d(x = M_emb[1,1], y = M_emb[1,2], M_emb[1,3], texts = "1")

## Do the Laplacian Eigenmap embedding
sklearn <- import("sklearn")
se <- sklearn$manifold$SpectralEmbedding(n_components = 2L, affinity = 'precomputed')
torus_le <- se$fit_transform(aff_matrix)

## Do the UMAP embedding
umap <- import("umap") ## installed from pip via umap-learn module
rot_k <- ceiling(log(nrow(M_emb))*3*sqrt(2)) ## Rule of thumb 
nn <- RANN::nn2(M_emb, query = M_emb, k = rot_k)
reducer <- umap$UMAP(n_components=2L, n_neighbors=as.integer(rot_k), 
                     min_dist=0.65, metric = "euclidean")
torus_umap <- reducer$fit_transform(M_emb)
# rgl::plot3d(torus_umap)
```

```{r}
## Torus Mapper
torus_mapper <- function(Z, overlap){
  function(){
    m <- Mapper:::MapperRef$new(as.matrix(M_emb))$
    use_cover(filter_values=Z, type="fixed rectangular", number_intervals=15L, percent_overlap=overlap)$
    use_distance_measure(measure = "euclidean")
    # use_clustering_algorithm(cl = "single", num_bins = 10L)$ # For analysis 
    m$clustering_algorithm <- function(X, idx, ...){ rep(1L, length(idx)) }
    m$compute_k_skeleton(k = 1L)$
      enable_multiscale()
  }
}

## Computes the mapper at the given discrete approximations. 'Exact' computes all poss
g_card <- list(10, 100, "exact")
torus_bench_le <- do.call(rbind, bench_mapper(torus_mapper(torus_le, 20), g_card))
torus_bench_umap <- do.call(rbind, bench_mapper(torus_mapper(torus_umap, 15), g_card))
saveRDS(torus_bench_le, file = "torus_bench_le.rds")
saveRDS(torus_bench_umap, file = "torus_bench_umap.rds")
```


Put together all the data to make the results plot
```{r}
## Data formatting for ggplot
format_data <- function(bench_res, type = c("set", "pair")){
  bench_cols <- switch(type, 
                       "set"=c("degree", "Naïve", "Indexed", "percent_overlap"), 
                       "pair" = c("degree", "Naïve", "Bounded", "Indexed", "percent_overlap"))
  bench_idx <- switch(type, "set"=2:3, "pair"=2:4)
  colnames(bench_res) <- bench_cols
  agg_res <- data.table::melt(bench_res, id.vars = c("degree", "percent_overlap"),
                              measure.vars=colnames(bench_res)[bench_idx])
  type <- sprintf("%s-%s", as.character(agg_res$variable), agg_res$degree)
  bench_data <- data.frame(type, agg_res$percent_overlap, agg_res$value)
  colnames(bench_data) <- c("approx_type", "percent_overlap", "counts")
  bench_data[["method"]] <- agg_res$variable
  bench_data[["approx_degree"]] <- agg_res$degree
  bench_data[["counts"]] <- log(bench_data$counts + 1) ## Add 1 to remove -Inf when 0 are compared at start
  return(bench_data)
}
normalize_overlap <- function(x, type=c("set", "pair")){
  cc <- ifelse(type == "set", 4, 5)
  if (ncol(x) == cc){ return(x) }
  else {
    normalize <- function(x){ (x - min(x))/(max(x) - min(x)) }
    return(cbind(x[, 1:cc], normalize(apply(x[, cc:ncol(x)], 1, sum))))
  }
}
diab_bench_ecc <- readRDS(file = "~/mapper/ignore/multiscale_demo/diab_bench_ecc.rds")
diab_bench_kde <- readRDS(file = "~/mapper/ignore/multiscale_demo/diab_bench_kde.rds")
wvs_bench_pca <- readRDS(file = "~/mapper/ignore/multiscale_demo/wvs_bench_pca.rds")
wvs_bench_ica <- readRDS(file = "~/mapper/ignore/multiscale_demo/wvs_bench_ica.rds")
torus_bench_le <- readRDS(file = "~/mapper/ignore/multiscale_demo/torus_bench_le.rds")
torus_bench_umap <- readRDS(file = "~/mapper/ignore/multiscale_demo/torus_bench_umap.rds")

## For d > 1, the sequence of overlaps is a vector, which in unamenable to plot as coordinate axis. 
## However, the linear combination of the values in the sequence is monotonically increasing.
type <- "pair"
# type <- "set"
type_idx <- ifelse(type=="set", 1L, 2L)
fpairs_diab_ecc <- format_data(normalize_overlap(do.call(rbind, diab_bench_ecc[, type_idx]), type), type)
fpairs_diab_kde <- format_data(normalize_overlap(do.call(rbind, diab_bench_kde[, type_idx]), type), type)
fpairs_wvs_pca <- format_data(normalize_overlap(do.call(rbind, wvs_bench_pca[, type_idx]), type), type)
fpairs_wvs_ica <- format_data(normalize_overlap(do.call(rbind, wvs_bench_ica[, type_idx]), type), type)
fpairs_torus_le <- format_data(normalize_overlap(do.call(rbind, torus_bench_le[, type_idx]), type), type)
fpairs_torus_umap <- format_data(normalize_overlap(do.call(rbind, torus_bench_umap[, type_idx]), type), type)


## Double legend line plot showing multiple groups 
make_line_plot <- function(x, title, sub_title){
  library("ggplot2")
  library("viridis")
  ## The fill aesthetic is dummy-group used to create two legends, because ggplot has a poorly designed interface.
  # y_breaks <- with(x, { seq(min(pairs), max(pairs)+0.5, by = 0.5) })
  # legend_title2 <- expression(paste(n^th, "-approximation:")
  legend_title2 <- "Approximation:"
  y_breaks <- ceiling(with(x, { seq(0, max(counts), length.out = 8) }))
  ggplot(data = x, aes(x=percent_overlap, group=approx_type, fill=approx_degree, color=method)) +
  geom_line(mapping = aes(y=counts, linetype=approx_degree, color=method), size=0.70) + 
  scale_linetype_manual(values=c("11", "twodash", "solid"), name=legend_title2) +
  scale_color_viridis(discrete = TRUE, name = "Method:", begin = 0, end = 0.8) +
  scale_y_continuous(breaks = y_breaks, position="left") + 
  theme_bw() +
  theme(legend.title = element_text(face="bold")) + 
  xlab("Percent Overlap") + ylab(label = "Cum. # operations (log-scale)") + 
  theme(axis.title=element_text(size=11))+
  ggtitle(title, subtitle = as.expression(parse(text=sprintf("italic('%s')", sub_title)))) +
  ## Legend Positions 
  # Bottom right corner 
  # theme(legend.position = c(0.98, 0.02), legend.direction = "vertical", legend.box = "horizontal",
  #       legend.title = element_text(size=6), legend.text = element_text(size=5), legend.key.size = unit(0.4, "cm"),
  #       legend.key.width=unit(0.8,"cm"), legend.box.margin = margin(0.025, 0.025, 0.025, 0.025, unit = "cm"),
  #       legend.spacing.x = unit(0.05, "cm"), legend.justification = c(1, 0),
  #       legend.box.background = element_rect(colour = "black"))
  # Top, horiozontal
  # theme(legend.position = "top", legend.direction = "horizontal", legend.box = "horizontal",
  #   legend.title = element_text(size=8), legend.text = element_text(size=6), legend.key.size = unit(0.02, "pt"),
  #   legend.key.width=unit(0.75,"pt"), legend.box.margin = margin(0.025, 0.025, 0.025, 0.025, unit = "pt"),
  #   legend.spacing = unit(0.25, "pt"), legend.justification = c(1, 1), legend.box.spacing = unit(0, "pt"),
  #   legend.box.background = element_rect(colour = "black"), legend.margin=margin(0.01, 0.03, 0.01, 0.03, unit = "pt"))
  # Bottom, horizontal 
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.box = "horizontal",
  legend.title = element_text(size=8), legend.text = element_text(size=6), 
  legend.key.width = unit(19.5, "pt"), 
  legend.box.margin = margin(1.5, 2.5, 1.5, 2.5, unit = "pt"),
  legend.spacing = unit(2.25, "pt"), legend.justification = "center", legend.box.spacing = unit(5, "pt"),
  legend.box.background = element_rect(colour = "black"), legend.margin=margin(1, 2, 1, 2, unit = "pt"))
}
data_names <- c("Miller-Reaves Chemical Diabetes data set", 
                "World Values Survey data set", 
                "Embedded Torus Synthetic data set")
g1 <- make_line_plot(fpairs_diab_ecc, title = data_names[1], sub_title = "KDE filter (1D)")
g2 <- make_line_plot(fpairs_diab_kde, title = data_names[1], sub_title = "Eccentricity filter (1D)")
g3 <- make_line_plot(fpairs_wvs_pca, title = data_names[2], sub_title = "Principle Components filter (3D)")
g4 <- make_line_plot(fpairs_wvs_ica, title = data_names[2], sub_title = "Independent Components filter (3D)")
g5 <- make_line_plot(fpairs_torus_le, title = data_names[3], sub_title = "Laplacian Eigenmap filter (2D)")
g6 <- make_line_plot(fpairs_torus_umap, title = data_names[3], sub_title = "UMAP filter (2D)")

## Save as PDF, per SIAM and ACM recommendations 
local({
  { page_width <- 390; page_height <- 592 } ## LaTeX inner body page width/height
  wh_ratio <- (page_height*0.33)/page_width ## To get figure aspect ratio to whatever percentage height of page
  res <- 3 ## to increase/decrease resolution
  { pt_width <- page_width*res; pt_height <- wh_ratio*pt_width } 
  pdf(file = paste0(type, "_results.pdf"), width = pt_width/72, height = pt_height/72, compress = TRUE)
  gridExtra::grid.arrange(g1, g3, g5, g2, g4, g6, nrow = 2, ncol = 3)
  dev.off()
})

## The fill aesthetic is dummy-group used to create two legends, because ggplot has a poorly designed interface.
# ggplot(data = bench_data, aes(x=percent_overlap, group=approx_type, fill=approx_degree, color=method)) +
#   geom_line(mapping = aes(y=pairs, linetype=approx_degree, color=method), size=0.90) + 
#   scale_linetype_manual(values=c("11", "twodash", "solid"), name=expression(paste(n^th, "-approximation"))) +
#   scale_color_viridis(discrete = TRUE, name = "Method", begin = 0, end = 0.8) +
#   scale_y_continuous(breaks = seq(min(vals), max(vals)+0.5, by = 0.5), position="right") + 
#   theme_bw() +
#   theme(legend.key.width = unit(1,"cm"), legend.position="left") + 
#   xlab("Percent Overlap") + ylab("Cumulative # Intersection Comparisons\n") +   
#   ggtitle("Computing mapper across scales", subtitle = expression(italic("Chemical Diabetes data set")))

## Legend horizontal at bottom 
# g1 + theme(legend.position = "bottom", legend.direction = "horizontal", legend.box = "vertical", legend.title = element_text(size=10), legend.text = element_text(size=8), legend.key.size = unit(0.6, "cm"), legend.key.width=unit(0.8,"cm"), legend.box.spacing = margin(0, 0, 0, 0, unit = "cm"), legend.spacing.y = unit(-0.3, "cm"))


```
```


```{r}

ls1 <- sapply(seq(length(mapper$cover$level_sets))-1L, function(i){
  multiscale$extract_level_set(i)
})


# wut <- FixedRectangularCover$new(filter_values = cbind(1:10, 1:10, 1:10))
# wut$number_intervals <- 3
# wut$percent_overlap <- 5
# bnds <- wut$level_set_bounds()
# idx <- expand.grid(1:3, 1:3, 1:3)
# sapply(1:nrow(idx), function(i) bnds[]
# rgl::points3d(bnds[,1:3])
```

## Smooth approximation of hausdorff distance
```{r}
c_mapper <- m_le
c_mapper <- m_umap


d_h.mds <- cmdscale(d = d_h, k = 3L, eig = FALSE)
```


To plot the torus
```{r}
## Plot with RGL + igraph 
am <- c_mapper$simplicial_complex$as_adjacency_matrix()
g <- igraph::graph_from_adjacency_matrix(am, mode = "undirected")
v_color_f1 <- bin_color(sapply(c_mapper$vertices, function(vids){ mean(c_mapper$X[vids, 1]) }))
v_color_f2 <- bin_color(sapply(c_mapper$vertices, function(vids){ mean(c_mapper$X[vids, 2]) }))

## Plot with RGL w/ first filter highlight
igraph::rglplot(g, layout = d_h.mds, vertex.size = 8, vertex.label = NA, 
                vertex.color = v_color_f1, 
                edge.color = "black", edge.width = 2)
rgl::axes3d()
rgl::title3d(main = NULL, sub = NULL, xlab = "mds1", ylab="mds2", zlab="mds3")

## Save RGL parameters to a list object
pp <- rgl::par3d(no.readonly=TRUE)
dput(pp, file="torus_view_le.R", control = "all")
  
## Plot with RGL w/ second filter 
pp <- dget("torus_view_le.R")
igraph::rglplot(g, layout = d_h.mds, vertex.size = 8, vertex.label = NA, 
                vertex.color = v_color_f2, 
                edge.color = "black", edge.width = 2)
rgl::axes3d()
rgl::title3d(main = NULL, sub = NULL, xlab = "mds1", ylab="mds2", zlab="mds3")
rgl::par3d(pp) ## view same orientation 
```



```{r}
g <- igraph::graph_from_adjacency_matrix(m$simplicial_complex$as_adjacency_matrix(), mode = "undirected", add.colnames = NA)

library(RColorBrewer)
diabetic_colors <- colorRamp(c("red", "white", "green"))(cbind(c(0, 0), c(1, 1)))
diabetic_colors
v_col <- sapply(names(m$vertices), function(vid) mean(as.integer(chemdiab$cc)[m$vertices[[vid]]]))+1

plot(g, vertex.color = palette()[v_col])
?igraph::plot.igraph

# gtour(flea[1:6], flea[[7]])
# gtour(olive[3:10], olive[[1]])
gtour(chemdiab[,1:5], as.integer(chemdiab$cc))
```

```{r}
# - Other examples - 
# ## Points unformly distributed on a 3-sphere
# sphere <- replicate(3, expr = rnorm(10000))
# sphere <- t(apply(sphere, 1, function(x) x/norm(matrix(x), type = "F")))
# 
# m_sphere <- MapperRef$new(X = sphere)
# m_sphere$
#   use_cover(filter_values = sphere[, c(1, 3)], typename = "fixed rectangular", 
#             number_intervals = 6L, percent_overlap = 20)$
#   use_distance_measure(measure = "euclidean")$
#   use_clustering_algorithm(cl = "single", num_bins = 10L)$
#   compute_k_skeleton(k = 1L)$
#   enable_multiscale() ## enables multiscale mapper
#   
# 
# el <- m_sphere$simplicial_complex$as_edge_list()
# ds <- Mapper::union_find(m_sphere$simplicial_complex$n_simplexes[[1]])
# invisible(apply(el, 1, function(x) ds$union(x[1], x[2])))
# H_0 <- length(unique(ds$connected_components())) ## first betti number

## double-torus embedded in R^3
# double_torus <- local({
#   { c <- 1; a <- 2 }
#   { u <- runif(1000, min = 0, max = 2*pi); v <- runif(1000, min = 0, max = 2*pi) }
#   cbind(
#     x = (c + a * cos(v)) * cos(u),
#     y = (c + a * cos(v)) * sin(u),
#     z = a * sin(v)
#   )
# })
```

```{r}

head(bench_res_ica)
bench_res_ica$percent_overlap <- apply(bench_res_ica[, 5:7], 1, function(x){ mean(x)})
bench_res_ica <- bench_res_ica[, c("resolution", "X1", "X2", "X3", "percent_overlap")]

make_line_plot(format_data(bench_res_ica), "1", "2")

outlier_idx <- detect_outliers(wv_ica$S, percentile = 0.95)
m_wvs <- MapperRef$
  new(as.matrix(WV_US.scaled)[-outlier_idx,])$
  use_cover(filter_values=wv_ica$S[-outlier_idx,], type="fixed rectangular", number_intervals=5L, percent_overlap=20)$
  use_distance_measure(measure = "euclidean")$
  use_clustering_algorithm(cl = "single", num_bins = 10L)$
  compute_k_skeleton(k = 1L)

g <- igraph::graph_from_adjacency_matrix(m_wvs$simplicial_complex$as_adjacency_matrix(), 
                                         mode = "undirected", add.colnames = NA)
normalize <- function(x, a=0, b=1){ (b-a)*((x - min(x))/(max(x) - min(x))) + a }
v_len <- sapply(m_wvs$vertices, length)
v_labels <- apply( cbind( names(m_wvs$vertices), v_len), 1, function(x) paste0(x, collapse = ":"))
v_size <- normalize(log(v_len), 10, 20)
v_color <- bin_color(log(v_len))
d_h <- hausdorff_distance(m_wvs)
dh_mds <- cmdscale(d_h, k = 2, eig = FALSE)
# v_coords <- dh_mds
v_coords <- igraph::layout_with_dh(g, coords = dh_mds, weight.node.dist = 1, 
                                   weight.node.edge.dist = 0.001*(1-edge_density(g)))
igraph::plot.igraph(g, 
    vertex.size = v_size, vertex.color = v_color, 
    vertex.label = v_labels, vertex.label.cex = 0.5, vertex.label.color = "black",
    edge.width = 0.75, layout = v_coords)

## Given two mappers and a set of coordinates corresponding to the vertices of the first mapper, 
## finds a set of coordinates for the vertices of that second mapper that approximates aligns to the 
## the first mapper
align_mappers <- function(coords, m1, m2){
  stopifnot(all(dim(m1$X) == dim(m2$X)))
  stopifnot(nrow(coords) == length(m1$vertices))
  n_vertex <- c(length(m1$vertices), length(m2$vertices))
  idx <- arrayInd(seq(prod(n_vertex)), .dim = n_vertex)
  min_dist <- apply(idx, 1, function(ii){
    v1 <- m1$vertices[[ii[1]]]
    v2 <- m2$vertices[[ii[2]]]
    mean(RANN::nn2(data = m1$X[v1,,drop=F], query = m2$X[v2,,drop=F], k = 1)$nn.dist)
  })
  min_v <- sapply(seq(n_vertex[[2]]), function(j){
    ii <- which(idx[, 2] == j)
    idx[ii[which.min(min_dist[ii])],1]
  })
  return(coords[min_v,])
}

wvs_mapper <- function(overlap){
  MapperRef$
  new(as.matrix(WV_US.scaled)[-outlier_idx,])$
  use_cover(filter_values=wv_ica$S[-outlier_idx,], type="fixed rectangular", number_intervals=5L, percent_overlap=overlap)$
  use_distance_measure(measure = "euclidean")$
  use_clustering_algorithm(cl = "single", num_bins = 10L)$
  compute_k_skeleton(k = 1L)
}
make_graph <- function(m){ 
  igraph::graph_from_adjacency_matrix(m$simplicial_complex$as_adjacency_matrix(), mode = "undirected") 
}

# igraph::layout_with_fr(make_graph(wv_m2), coords = coord2, start.temp = 0.01, niter = 10)
test_g <- seq(2.5, 25, by = 2.5)
wv_mappers <- lapply(test_g, function(g){
  wvs_mapper(g)
})
connected_mapper <- tail(wv_mappers, 1)[[1]]
d_h <- Mapper::hausdorff_distance(connected_mapper)
base_coords <- apply(cmdscale(d_h, k = 2, eig = FALSE), 2, normalize)


base_grapher <- connected_mapper$as_grapher(FALSE)
base_grapher$nodes$x <- base_coords[, 1]
base_grapher$nodes$y <- base_coords[, 2]

n_nodes <- connected_mapper$simplicial_complex$n_simplexes[[1]]
n_edges <- connected_mapper$simplicial_complex$n_simplexes[[2]]
v_color <- bin_color(log(sapply(connected_mapper$vertices, length)))
G <- grapher(base_grapher) %>% 
  setLinkColor(color = rep(rgb(0, 0, 0), n_edges)) %>% 
  setNodeColor(color = v_color) %>% 
  enableForce()


library(shiny)
ui <- pageWithSidebar(
  headerPanel("Simple shiny app with Grapher"),
  sidebarPanel(
    actionButton(inputId = "toggle_force", label = "Toggle force"),
    actionButton(inputId = "get_network", label = "Get network")
  ),
  ## The main canvas container holding the grapher
  mainPanel( grapherOutput(outputId = "grapher", height = "100vh") )
)

server <- function(input, output, session) {
  ## Pass the htmlwidget to the canavs container via 'renderGrapher'
  enable_force <<- TRUE
  output$grapher <- renderGrapher({ G })
  observeEvent(input$get_network, {
    session <- shiny::getDefaultReactiveDomain()
    session$sendCustomMessage("grapher:getNetwork", list(id="grapher"))
  })
  observeEvent(input$network, {
    req(input$network)
    network <<- input$network
  })
  observeEvent(input$toggle_force, {
    req(input$toggle_force)
    if (enable_force){
      enableForce("grapher")
      enable_force <<- FALSE
    } else {
      disableForce("grapher")
      enable_force <<- TRUE
    }
  })
}

shinyApp(ui, server)

nice_coords <- do.call(rbind, lapply(network$nodes, function(v) with(v, c(x, y))))

all_coords <- lapply(wv_mappers, function(m){
  align_mappers(nice_coords, connected_mapper, m)
})


g <- make_graph(wv_mappers[[1]])
x <- igraph::layout.davidson.harel(g, coords = all_coords[[1]], 
                            maxiter = 10, fineiter = 10, cool.fact = 0.75, weight.node.dist = 1, 
                            weight.border = 10, weight.edge.lengths = edge_density(g)/10, 
                            weight.edge.crossings = 1 - sqrt(edge_density(g)), 
                            weight.node.edge.dist = 0.2 * (1 - edge_density(g)))
plot(all_coords[[1]], xlim = range(c(all_coords[[1]][, 1], x[, 1])), ylim = range(c(all_coords[[1]][, 2], x[, 2])))
points(x, col = "blue")

# all_g <- lapply(wv_mappers, make_graph)

plot_mapper_igraph <- function(m, v_coords, ...){
  normalize <- function(x, a=0, b=1){ (b-a)*((x - min(x))/(max(x) - min(x))) + a }
  v_len <- sapply(m$vertices, length)
  v_labels <- apply( cbind( names(m$vertices), v_len), 1, function(x) paste0(x, collapse = ":"))
  v_size <- normalize(log(v_len), 4, 12)
  v_color <- bin_color(log(v_len))
  igraph::plot.igraph(g, 
      vertex.size = v_size, vertex.color = v_color, 
      vertex.label = v_labels, vertex.label.cex = 0.5, vertex.label.color = "black",
      edge.width = 0.75, layout = v_coords, ...)
}

layout(matrix(seq(4), nrow = 2, ncol = 2))
par(mar = c(0, 0, 0, 0))
invisible(lapply(seq(1), function(i){
  perturbed_coords <- igraph::layout_with_fr(
    make_graph(wv_mappers[[i]]), coords = all_coords[[i]], 
    niter = 100, start.temp = 45
  )
  plot_mapper_igraph(wv_mappers[[i]], perturbed_coords)
}))



igraph::layout_with_fr(graph = g2, coords = coord2)

plot(g1, layout = base_coords)

wut <- grapher::getDefaultJsonConfig(network = 

grapher::grapher(wut) %>% 
  grapher::setNodes()
#   plot(coords )
#   text(coords, labels = 1:nrow(coords), cex = 0.4, pos = 3)
#   points(coords[min_v,], col = "green", pch = 20, cex = 0.4)
#   text(coords[min_v,], labels = 1:nrow(coords[min_v,]), cex = 0.4, pos = 3, col = "green")
# arrayInd(seq(prod(rep(3, 3))), .dim = 3)
#   idx <- combn(n_vertex, 2)
#   d_h.vec <- apply(idx, 2, function(ii){
#     v1 <- m$vertices[[ii[1]]]
#     v2 <- m$vertices[[ii[2]]]
#     xy <- mean(RANN::nn2(data = m$X[v1,,drop=F], query = m$X[v2,,drop=F], k = 1)$nn.dist)
#     yx <- mean(RANN::nn2(data = m$X[v2,,drop=F], query = m$X[v1,,drop=F], k = 1)$nn.dist)
#     max(xy, yx)
#   })
# }




# agg_res <- data.table::melt(tmp, id.vars = c("degree", "percent_overlap"), measure.vars=colnames(tmp)[2:4])
linetypes <- c("dotted", "dashed", "solid")
library("ggplot2")
ggplot2::ggplot(data = tmp, aes(x=percent_overlap, group=degree)) +
  geom_line(mapping = aes(y=log(`Brute Force`), linetype=degree), color="red", size=0.8) +    
  geom_line(mapping = aes(y=log(`Bounded Approach`), linetype=degree), color="blue", size=0.8) +
  geom_line(mapping = aes(y=log(`Indexed Approach`), linetype=degree), color="green", size=0.8)+
  scale_linetype_manual(values=linetypes)+
  theme(legend.position="right")
```


```{r}
library(shiny)
library(ggvis)
library(tourr)
gtour <- function(data, groups) {
  aps <- 2
  fps <- 30
  mat <- rescale(as.matrix(data))
  tour <- new_tour(mat, grand_tour(), NULL)
  start <- tour(0)
  proj_data <- reactive({
    invalidateLater(1000 / fps, NULL);
    step <- tour(aps / fps)
    data.frame(center(mat %*% step$proj), groups = groups)
  })
  proj_data %>% ggvis(~X1, ~X2, fill = ~groups) %>%
    layer_points() %>%
    scale_numeric("x", domain = c(-1, 1)) %>%
    scale_numeric("y", domain = c(-1, 1)) %>%
    add_axis("x", title = "") %>% add_axis("y", title = "") %>%
    set_options(duration = 0)
}

```



```{r}
multiscale_le <- m_le$.__enclos_env__[["private"]][[".multiscale"]]

np <- import("numpy")
sklearn <- import("sklearn")
scaled.laplacian <- apply(graph_laplacian[idx, idx], 2, scale)
aff_matrix <- r_to_py(as.matrix(torus_kernel)[1:500, 1:500], TRUE)
spec_emb <- sklearn$manifold$SpectralEmbedding(n_components=2, affinity = 'precomputed')
wut <- spec_emb$fit(aff_matrix)
wut <- spec_emb$affinity

le_emb <- sklearn$manifold$spectral_embedding(adjacency = aff_matrix, n_components = 2, norm_laplacian = TRUE)
# graph_laplacian <- igraph::graph.laplacian(igraph::make_full_graph(n), normalized = TRUE, weights = torus_kernel)
# saveRDS(graph_laplacian, file = "graph_laplacian.rds")
graph_laplacian <- readRDS(file = "graph_laplacian.rds")

## Check Matrix::isSymmetric(graph_laplacian) == TRUE
idx <- sample(1:n, replace = FALSE, size = 500L)
gl_eig_vec <- RSpectra::eigs_sym(graph_laplacian[idx, idx], k = 2L, which = "LM")

gl_eig_vec <- RSpectra::eigs_sym(graph_laplacian[idx, idx], k = 2L, which = "LM")

gl_eig <- eigen(graph_laplacian)
?Matrix::solve
heated_graph <- igraph::graph_from_adjacency_matrix(g, mode = "undirected", weighted = TRUE)

D <- diag(apply(g, 2, sum), nrow(g), ncol(g))
L <- D - g
eigen(L)
```

