---
author: "Matt Piekenbrock"
title: "Using a custom covering method"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{custom_cover}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

Perhaps one of the most important aspects of parameterizing Mapper is the choice of the cover.
Given data $X$ and a continuous map $f : X \to Z$, the cover (often denoted as $\mathcal{U} = \{U_\alpha\}_{\alpha \in A}$) is constructed over the values of $Z$. Since $f$ is continuous, $\mathcal{U}$ also forms a cover of $X$. That is, the sets 

$$\{ f^{-1}(U_\alpha), \alpha \in A \}$$
form an open cover over $X$, i.e. $X \subseteq \bigcup\limits_{\alpha \in A} f^{-1}(U_\alpha)$. By applying a _partial clustering_ to each of these open sets in $X$, $f^{-1}(U_\alpha)$ decomposes into a set of path-connected components. These form the $0$-simplices of the resulting _mapper_ construction. 

The `Mapper` package theoretically can support any covering strategy, so long as a few requirements are satisfied. All covers can be constructed either directly via their R6 generators, or more succinctly using the \code{$use_cover} member function associated with a \code{MapperRef} object. See \code{?MapperRef} for more details. A few frequently used methods are included in the package. Their typenames, generators, and parameters can be listed via: 
```{r}
Mapper::covers_available()
```


## Making a custom cover 

To use a custom covering strategy, a new `R6` class must be defined satisfying the following requirements:

1. The class inherits from the `CoverRef` class. 

2. The class overrides the `construct_cover` function. 

3. Upon calling `construct_cover`, the class populates the `level_sets` member with a named list whose names are uniquely indexed by the `index_set` member, and whose values are integer indices that form a valid cover. 

Below is an example of a simplified version of the source code used to create the `BallCover` generator.  

```{r}
BallCoverEx <- R6::R6Class(
  classname = "BallCoverEx",
  inherit = CoverRef,
  public = list(
    epsilon=NA,   ## cover parameter    
    initialize = function(epsilon){
      super$initialize(typename="ball")
      self$epsilon <- epsilon
    }
  )
)
```
First, a generator is defined that satisfies (1). The public fields represent the parameters of the cover. Note that the initializer must be overridden so that the filter values along with the unique 'typename' of the cover are first passed to the parent `CoverRef` initializer. When that happens, a few checks are performed, and the following private fields become available to the derived class via the `private` namespace.  

__private`$`__

- `__.level_sets:__` list of data indices that intersect the open sets

- `__.index_set:__` character vector of keys that uniquely index the level sets

- `__.typename:__` string identifier of the covering method

Additional information can be found in the documentation of the base class, see `?CoverRef` for more details. 

The only other required function to override is the `construct_cover` method. Below is an example that uses the `RANN` package to construct a cover. 
```{r}
BallCoverEx$set("public", "construct_cover", function(filter, index=NULL){
  ## Get filter values 
  fv <- filter()
  f_dim <- ncol(fv)
  f_size <- nrow(fv)
  
  ## Construct the balls
  ball_cover  <- RANN::nn2(fv, query = fv, searchtype = "radius", radius = self$epsilon)
  
  ## Union them together 
  ds <- union_find(f_size)
  apply(ball_cover$nn.idx, 1, function(idx){
    connected_idx <- idx[idx != 0] - 1L
    if (length(connected_idx) > 0){ ds$union_all(connected_idx) }
  })
  
  ## Construct the intersections between the open sets and the data
  cc <- ds$connected_components()
  self$index_set <- as.character(unique(cc))
  ls <- lapply(self$index_set, function(idx){ which(cc == as.integer(idx)) })
  self$level_sets <- structure(ls, names=self$index_set)
  
  ## If index is requested, return the preimages at that index, 
  ## otherwise return the instance invisibly
  if (!missing(index)){ return(self$level_sets[[index]]) }
  invisible(self)
})
```

A [disjoint-set data structure](https://en.wikipedia.org/wiki/Disjoint-set_data_structure) is used to track the intersections between each ball (see `?union_find` for usage details). 
Note how the index set uniquely indexes the open sets. Although in this case the balls don't necessarily represent the classical notion of a "level set", the member is named that way due to their relaxed interpretation as such in the other covers. 

## Making a better custom cover

The above example is all one needs to add use a cover not included in the package. However, there are other methods or extensions that may be worth implementing. For example, the above example cover isn't type-safe. A better example would use [active bindings](https://adv-r.hadley.nz/r6.html#active-fields) to make fields which can be assigned to like regular member fields, but also type-check on assignment. 

Another important method that can be overridden is the `$level_sets_to_compare` member, which is inherited by the base class. This method controls which of the open sets to compare (pairwise) in computing the $1$-skeleton. By default, it returns all pairwise combinations of indices, which can obviously be greatly improved depending on how the cover is parameterized!  

Once you have a cover generator, you can check if its compatible with the package using the `$validate` function. If it does, it should silently return. 

## Contributing

If you do design a new covering method not used in the package, consider submitting a [pull request](https://github.com/peekxc/mapper/pulls) incase others might find it useful! 

