---
title: "Using a custom covering method"
output: html_notebook
---

## Introduction
Perhaps one of the most important aspects of parameterizing Mapper is the choice of the cover.
Given data $X$ and a continuous map $f : X \to Z$, the cover (often denoted as $\mathcal{U} = \{U_\alpha\}_{\alpha \in A}$) is constructed over the values of $Z$. Since $f$ is continuous, $\mathcal{U}$ also forms a cover of $X$. That is, the sets 

$$\{ f^{-1}(U_\alpha), \alpha \in A \}$$
form an open cover over $X$, i.e. $X \subseteq \bigcup\limits_{\alpha \in A} f^{-1}(U_\alpha)$. By applying a _partial clustering_ to each of these open sets in $X$, $f^{-1}(U_\alpha)$ decomposes into a set of path-connected components. These form the $0$-simplices of the resulting _mapper_ construction. 

The `Mapper` package works with any covering strategy, so long as a few requirements are satisfied. All covers can be constructed either directly via their R6 generators, or more succinctly using the \code{$use_cover} member function associated with a \code{MapperRef} object. See \code{?MapperRef} for more details. A few frequently used methods are included in the package. Their typenames, generators, and parameters can be listed via: 
```{r}
Mapper::covers_available()
```


## Making a custom cover 

To use a custom covering strategy, a new `R6` class must be defined satisfying the following requirements:
1. The class inherits from the `CoverRef` class. 
2. The class overrides the `construct_cover` function. 
3. Upon calling `construct_cover`, the class populates the `level_sets` member with a named list whose names are uniquely indexed by the `index_set` member, and whose values are integer indices that form a valid cover. 

Below is an example of a simplified version of the source code used to create the `AdaptiveCover` generator. First, a generator is defined that satisfies (1). The public fields represent the parameters of the cover.  

```{r}
## Parameters to make the cover
params <- list(number_intervals=NA, percent_overlap=NA, quantile_method=NA)

## The generator
# AdaptiveCover_ex <- R6::R6Class(
#   classname = "AdaptiveCover_ex",
#   inherit = CoverRef,
#   public = append(params, 
#     list(initialize = function(filter_values, number_intervals, percent_overlap, quantile_method){
#         super$initialize(filter_values, typename="adaptive cover example")
#         self$number_intervals <- number_intervals
#         self$percent_overlap <- percent_overlap 
#         self$quantile_method <- quantile_method
#     })
#   )
# )
```

In the example above, 

```{r}
CoverRef
```


#' \itemize{
#'  \item{.level_sets}{names list of indices in the original data set to cluster over.}
#'  \item{.index_set}{character vector of keys that uniquely index the level sets.}
#'  \item{.filter_dim}{constant representing the filter dimension.}
#'  \item{.filter_size}{constant representing the number of points in the filter space.}
#'  \item{.typename}{string identifier of the covering method.}
#' }
