---
title: "R Notebook"
output: html_notebook
---


```{r}
library(Mapper)
X <- replicate(n = 3, expr = rnorm(1000))
f_x <- apply(X, 1, function(x){
   max(abs(x - X[which.min(X[,1]),]))
})

m <- MapperRef$new(as.matrix(X))$
 use_cover(filter_values=matrix(f_x), type="fixed rectangular", number_intervals=20L, percent_overlap=0)$
 use_distance_measure(measure = "euclidean")
m$clustering_algorithm <- function(X, idx){ rep(1L, length(idx)) }
m$compute_k_skeleton(k = 1L)$
  enable_multiscale()


## Basic profiling
profvis::profvis({
  invisible(sapply(seq(0.005, 50, by = 0.005), function(g_i){
      m$update_mapper(percent_overlap = c(g_i))
   })) 
})

```


```{r}
library("Mapper")
ls_pairs <- t(combn(4, 2))
nodes <- list(1:5, 2:6, 3:7, 4:8)
ls_node_map <- list(1:2, 2:3, 3:4, c(1, 4)) 

n_vertices <- 5L
stree <- Mapper::simplex_tree()
stree_ptr <- stree$as_XPtr()
Mapper:::build_1_skeleton(ls_pairs = ls_pairs, nodes = nodes, ls_node_map = ls_node_map, stree = stree_ptr)
```

```{Rcpp}

// [[Rcpp::plugins(cpp11)]]

#include <Rcpp.h> 
using namespace Rcpp;

// Simple equivalent to unlist
// [[Rcpp::export]]
IntegerVector combine_int(const List& list)
{
   std::size_t n = list.size();
   std::size_t total_length = 0;
   for (std::size_t i = 0; i < n; ++i) total_length += Rf_length(list[i]);
   IntegerVector output = no_init(total_length);
   std::size_t index = 0;
   for (std::size_t i = 0; i < n; ++i)
   {
      std::vector<int> el = list[i];
      std::copy(el.begin(), el.end(), output.begin() + index);
      index += el.size();
   }
   return output;
}

// [[Rcpp::export]]
List combine_lst(){
  std::vector< std::vector<int> > vec = std::vector< std::vector<int> >();
  for (int i = 0; i < 5; ++i){
    vec.push_back(std::vector<int>(10, 0));
  }
  return(wrap(&vec.at(1)));
}

// [[Rcpp::export]]
IntegerVector list_subset(const List& x, IntegerVector idx){
//  List lst_sub = x[idx];
  return(combine_int(x[idx]));
//  return(x[idx]);
}

```


```{r}
lst <- list(1:10, 11:20, 21:30)
list_subset(lst, as.integer(1:2))
list_subset(lst, as.integer(1:2))
```
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

